---
title: "De-Criminalizing Suicide: Analysis"
author: "Nishtha Arora"
output: html_document
date: "2023-03-19"
---
# COMPARISON WITH DEVELOPMENT STATUS. 2008-2017. CAN BE TILL 2019. ALL COUNTRIES

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Research statements

- India’s Mental Healthcare Act 2017 is not enough to help decriminalize suicide. There are still loopholes as to what is the limit of severe stress when deciding the case of attempted suicide.

- Importance and limitations of suicide registry database in providing data and information for suicide intervention programmes. As one of the effects of criminalizing suicide is the lack of data and resources available for meaningful reporting and evaluation of suicide cases.

- Actions and solutions on what should countries like Msia prepare in their move to decriminalize suicide. 
# Data sources 
1. Using data from 2008-2017 (10 years), worldwide suicide rates- 
https://ourworldindata.org/suicide

2. Using HDI data for High, Very high and medium development country recognition.

<!-- Using -->
<!-- https://ourworldindata.org/world-population-growth -->

# Analysis

```{r libraries}
library(readr)
library(tidyverse)
library(readxl)
```

# Countries development status

```{r cleaning}
read_data <- function(x,y,z){
  read_excel("data/HDIstatus.xlsx", rang = x,col_names = FALSE) %>% 
    rename(HDI_rank_2021 = ...1, 
           Country = ...2, 
           HDI_Value = ...3, 
           Life_expectancy = ...5, 
           Expected_years_of_schooling = ...7, 
           Mean_years_of_schooling = ...9, 
           GNI_per_capita = ...11, 
           GNI_rank_minus_HDI_rank = ...13, 
           HDI_rank_2020 = ...15)} %>% 
  select(HDI_rank_2021, Country, HDI_Value,
         Life_expectancy, Expected_years_of_schooling, 
         Mean_years_of_schooling, GNI_per_capita, 
         GNI_rank_minus_HDI_rank, HDI_rank_2020) %>% 
  mutate(y = z )


VERY_HIGH_HUMAN_DEVELOPMENT <- 
  read_data('A9:O74',Degree_of_Human_Development, 'VERY HIGH HUMAN DEVELOPMENT')

HIGH_HUMAN_DEVELOPMENT <- 
  read_data('A76:O128', Degree_of_Human_Development, 'HIGH HUMAN DEVELOPMENT' ) 

MEDIUM_HUMAN_DEVELOPMENT <- 
  read_data('A130:O166', Degree_of_Human_Development, 'MEDIUM HUMAN DEVELOPMENT' ) 

LOW_HUMAN_DEVELOPMENT <- 
  read_data('A168:O200', Degree_of_Human_Development, 'LOW HUMAN DEVELOPMENT' )

HDI <- bind_rows(VERY_HIGH_HUMAN_DEVELOPMENT,HIGH_HUMAN_DEVELOPMENT,MEDIUM_HUMAN_DEVELOPMENT,LOW_HUMAN_DEVELOPMENT)

write_csv(HDI, "data/HDIStatus2.csv")
```

```{r}
data3 <- read_csv("data/HDIStatus2.csv") %>% 
  select(y, Country) %>% 
  filter( y %in% c("VERY HIGH HUMAN DEVELOPMENT", "HIGH HUMAN DEVELOPMENT", 
                   "MEDIUM HUMAN DEVELOPMENT")) %>% 
  rename("Entity"= Country)
```

# Suicide Rates worldwide 

```{r reading data}
data2 <- read_csv("data/suicide-death-rates.csv") 
```

```{r}
filtered <- data2 %>% filter(between(Year, ('2008'),('2017'))) %>% 
  group_by(Entity) %>% 
  summarise(avg_rate=mean(`Deaths - Self-harm - Sex: Both - Age: Age-standardized (Rate)`))
```

```{r}
join <- full_join(filtered, data3) %>% 
  arrange(desc(avg_rate)) %>% na.omit() 
```

```{r}
one <- join %>% filter(y== "HIGH HUMAN DEVELOPMENT") %>% head(10) %>% arrange()
two <- join %>% filter(y== "VERY HIGH HUMAN DEVELOPMENT") %>% head(10) %>% arrange()
three <- join %>% filter(y== "MEDIUM HUMAN DEVELOPMENT")%>% head(10)%>% arrange()
```

# Visualization for top 10 countries (in terms of highest average suicide rates of years 2008-2017) of each selected category.

```{r}
plot1 <- join %>% head(30)
ggplot(plot1, aes(x=Entity, y=avg_rate, fill=y))+
  geom_col()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  labs(title="Top 10 countries with maximum suicides",
         x="Country",
      y="Rate")
```



# Suicides in India due to Mental Health? How many people have mental health issues?
How many are getting treated for it?

# Which countries de-criminalize it, and which dont. So if they do, why not india at this number?

# 





https://www.kaggle.com/datasets/gauravdahiya/all-india-distribution-of-suicides-during-2018?resource=download



