---
title: "Article 2"
author: "Nishtha Arora"
date: "2023-04-04"
output: html_document
---

Suicide prevention in Bangladesh: The current state and the way forward

Bangladesh Mental Health Act 2018

Among the eight South Asian countries, suicide (including attempt) is currently a crime in Bangladesh only, but has been decriminalized in Bhutan, Nepal, India, Afghanistan, Maldives, and Sri Lanka. Pakistan decriminalised suicide in December 2022.

Data source- https://ourworldindata.org/suicide

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Bangladesh suicide rate for 2019 was 3.70, a 2.78% increase from 2018.

```{r}
death_rate_from_suicides_gho <- read_csv("data/branch2/death-rate-from-suicides-gho.csv")
```

```{r}
Bangladesh <- death_rate_from_suicides_gho %>% filter(Entity == "Bangladesh")
```

```{r}
colnames(Bangladesh)
```

```{r}
str(Bangladesh)
```


```{r}
plot <- ggplot(Bangladesh, aes(x=Year, y=`Age-standardized suicide rate - Sex: both sexes` ))+
  geom_line(color="steelblue")+
  geom_point()+
  ylab("Suicide Rate")+
  ggtitle("Bangladesh")+
theme_classic()

```

Suicides started increasin from 2016

In 2016, the government of Bangladesh introduced a pilot project to strengthen the civil registration and vital statistics system and generate cause of death data in Kaliganj Upazila. Community-based health workers were trained to notify births and deaths to the civil registrar, and to conduct verbal autopsy interviews with family members of a deceased person

source- https://pubmed.ncbi.nlm.nih.gov/31474777/


