---
title: "De-Criminalizing Suicide:  Initial Analysis"
author: "Nishtha Arora"
output: html_document
date: "2023-03-19"
---
# Initial Analysis 1

## COMPARISON WITH DEVELOPMENT STATUS. 2008-2019.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width=12, fig.height=8)
```

### Data sources 
1. Using data from 2008-2019 (12 years), worldwide suicide rates- 
https://ourworldindata.org/suicide

2. Using HDI data for High, Very high and medium development country recognition- http://hdr.undp.org/en/composite/HDI


```{r libraries}
library(readr)
library(tidyverse)
library(readxl)
```

```{r cleaning}
read_data <- function(x,y,z){
  read_excel("data/InitialA1/HDIstatus.xlsx", rang = x,col_names = FALSE) %>% 
    rename(HDI_rank_2021 = ...1, 
           Country = ...2, 
           HDI_Value = ...3, 
           Life_expectancy = ...5, 
           Expected_years_of_schooling = ...7, 
           Mean_years_of_schooling = ...9, 
           GNI_per_capita = ...11, 
           GNI_rank_minus_HDI_rank = ...13, 
           HDI_rank_2020 = ...15)} %>% 
  select(HDI_rank_2021, Country, HDI_Value,
         Life_expectancy, Expected_years_of_schooling, 
         Mean_years_of_schooling, GNI_per_capita, 
         GNI_rank_minus_HDI_rank, HDI_rank_2020) %>% 
  mutate(y = z )


VERY_HIGH_HUMAN_DEVELOPMENT <- 
  read_data('A9:O74',Degree_of_Human_Development, 'VERY HIGH HUMAN DEVELOPMENT')

HIGH_HUMAN_DEVELOPMENT <- 
  read_data('A76:O128', Degree_of_Human_Development, 'HIGH HUMAN DEVELOPMENT' ) 

MEDIUM_HUMAN_DEVELOPMENT <- 
  read_data('A130:O166', Degree_of_Human_Development, 'MEDIUM HUMAN DEVELOPMENT' ) 

LOW_HUMAN_DEVELOPMENT <- 
  read_data('A168:O200', Degree_of_Human_Development, 'LOW HUMAN DEVELOPMENT' )

HDI <- bind_rows(VERY_HIGH_HUMAN_DEVELOPMENT,HIGH_HUMAN_DEVELOPMENT,MEDIUM_HUMAN_DEVELOPMENT,LOW_HUMAN_DEVELOPMENT)

write_csv(HDI, "data/InitialA1/HDIStatus2.csv")
```

```{r}
data3 <- read_csv("data/InitialA1/HDIStatus2.csv") %>% 
  rename("Entity"= Country)
```

```{r reading data}
data2 <- read_csv("data/InitialA1/suicide-death-rates.csv")
```

```{r}
# setdiff(data2$Entity, data3$Entity)

change <- data2 %>%
  mutate(Entity = recode(Entity,
    "American Samoa" = "Samoa",
    "Bolivia" = "Bolivia (Plurinational State of)",
    "Brunei"= "Brunei Darussalam",
    "Cape Verde"= "Cabo Verde",
    "Cote d'Ivoire"="Côte d'Ivoire",
    "Democratic Republic of Congo"= "Congo (Democratic Republic of the)",
    "Eswatini"= "Eswatini (Kingdom of)",
    "Iran"= "Iran (Islamic Republic of)",
    "Laos"= "Lao People's Democratic Republic",
    "Micronesia (country)"= "Micronesia (Federated States of)",
    "North Korea"= "Korea (Republic of)",
    "Northern Ireland"= "Ireland",
    "Palestine"= "Palestine, State of",
    "Russia"= "Russian Federation",
    "South Sudan"= "Sudan",
    "Syria"= "Syrian Arab Republic",
    "Timor"= "Timor-Leste",
    "Venezuela"= "Venezuela (Bolivarian Republic of)",
    "Vietnam"= "Viet Nam"
    ))
```


```{r}
filtered <- change %>% filter(Year > 2007) %>% 
  group_by(Entity) %>% 
  summarise(avg_rate=mean(`Deaths - Self-harm - Sex: Both - Age: Age-standardized (Rate)`))
```


```{r}
join <- full_join(filtered, data3) %>% 
  arrange(desc(avg_rate)) %>% na.omit() %>% 
  rename(`Development status` = y) %>% 
  mutate(avg_rate= round(avg_rate, digits=2))

```

### Visualization for top 30 countries with maximum suicide rates

```{r}
plot1 <- join %>% head(30)
ggplot(plot1, aes(x=reorder(Entity, -avg_rate), y=avg_rate, fill=`Development status`))+
  geom_col()+
geom_text(aes(label = avg_rate), vjust = 2, colour = "white", size=3)+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  labs(title="Top 30 countries with maximum suicides, 2008-2019",
         x=substitute(paste(bold("Country"))),
      y=substitute(paste(bold("Suicide rate per 100,000 people"))))+
  scale_fill_manual(values = c("#A0331C", "#1C56A0", "#4B902F", "#635A61"))+
   theme(plot.title = element_text(face="bold"))
```

# Initial Analysis 2

## India's Mental Health Act 2017

### Suicide Rates & reported mental health issues before and after 2017

Data source: https://data.worldbank.org/indicator/SH.STA.SUIC.P5?locations=IN
https://ourworldindata.org/mental-health

```{r}
library(ggthemes)
library(reshape2)

Anyl2 <- data2 %>% filter(Entity == "India") %>% 
  filter(Year >1999)

data4 <- read_csv("data/InitialA1/share-with-mental-and-substance-disorders.csv")%>% 
  filter(Entity == "India")  %>% 
  filter(Year >1999)

change_join <- full_join(Anyl2, data4, by = c("Year", "Code", "Entity")) %>%  
  select(-c(Entity, Code))
final_join <-  melt(change_join, id.vars = "Year") %>%
  mutate(variable = recode(variable,
    "Deaths - Self-harm - Sex: Both - Age: Age-standardized (Rate)" = "Suicide",
    "Prevalence - Mental disorders - Sex: Both - Age: Age-standardized (Percent)"= "Mental Health disorder"
    ))
```


```{r}
ggplot(final_join, aes(x=Year, y=value, col=variable))+
  geom_line()+
  ylab(substitute(paste(bold("Rate per 100,000 people"))))+
  xlab(substitute(paste(bold(Year))))+
  geom_point(color="brown", size=1.5)+
  ggtitle("Time-series for India, 2000-2019")+
 theme_economist()+
   theme(plot.title = element_text(size=14, lineheight=.8))+

  geom_vline(xintercept = 2017, linetype = "dashed")+
  scale_x_continuous(breaks = seq(from= 2000,to=2019, by =1 ))+
scale_color_manual(values=c('Brown','Purple', 'Black'))+
  geom_text(aes(x = 2005 , y = 18.5, label = "Suicide Rate", colour = "Black")) +
geom_text(aes(x = 2010, y = 14.2, label = "Mental Disorder rate", colour= "Black"))+
  theme(legend.position = "none")
```








