---
title: "Policing the police"
author: "Nishtha Arora"
date: "2023-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading_libraries}
library(readxl)
library(readr)
library(readxl)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(scales)
```

# Visualization 1 & 2

## HIGH CORRUPTION COUNTS GLOBALLY

### Data source:

- UNODC: https://dataunodc.un.org/dp-crime-corruption-offences

```{r daya_unodc_plot1,2}
corruption_UNODC <-
  read_excel("data/policingthepolice/data_cts_corruption_and_economic_crime.xlsx") |>
  slice(-c(1, 2)) |>
  set_names(c(
    "Iso3_code",
    "Country",
    "Region",
    "Subregion",
    "Indicator",
    "Dimension",
    "Category",
    "Sex",
    "Age",
    "Year",
    "Unit of measurement",
    "VALUE",
    "Source"
  ))

corruption_UNODC_clean <- corruption_UNODC |>
  filter(`Unit of measurement` == "Counts") |>
  select(Iso3_code, Country, Category, VALUE, Year) |>
  filter(Category == "Corruption") |>
  mutate(VALUE = as.numeric(VALUE))

plot_global <- corruption_UNODC_clean |>
  group_by(Country) |>
  summarise(VALUE = mean(VALUE)) |>
  arrange(desc(VALUE)) |>
  head(10) |>
  mutate(ToHighlight = ifelse(VALUE > 30000, "yes", "no")) |>
  ggplot(aes(reorder(x = Country, -VALUE), y = VALUE, fill = ToHighlight)) +
  geom_col() +
  theme(legend.position = "none") +
  theme_dark() +
  ggtitle("Average corruption count, 2013-2021") +
  xlab("Country") +
  ylab("Corruption Count") +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("yes" = "#57AFD5", "no" = "lightblue"))

plot_global

plot_usa <- corruption_UNODC_clean |>
  filter(Country == "United States of America") |>
  mutate(
    Year = as.Date(as.character(Year), format = "%Y"),
    Year = year(Year)
  ) |>
  ggplot(aes(x = Year, y = VALUE)) +
  geom_col(fill = "lightblue", width = 0.4) +
  geom_line(size = 1, colour = "#811B0F") +
  ggtitle("Corruption Count: USA") +
  ylab("Corruption Count") +
  theme_dark() +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  )

plot_usa
```

# Visualization 3

## PERCEPTION OF CORRUPTION SURVEY, 2013-2017

### Data source:

- OWID (Our World in Data): https://ourworldindata.org/corruption

```{r data_owid_plot3}
OWID_police <-
  read_csv("data/policingthepolice/perception-of-corruption-by-institution.csv") |>
  group_by(Year) |>
  summarise_if(is.numeric, mean, na.rm = TRUE) |>
  pivot_longer(cols = -Year, names_to = "Institution", values_to = "Count") |>
  ungroup()

# sum counts across all years
OWID_police_allyears <-
  OWID_police |>
  group_by(Institution) |>
  summarise(Count = sum(Count, na.rm = TRUE)) |>
  mutate(Institution = fct_reorder(Institution, Count, .desc = TRUE))

ggplot(OWID_police_allyears, aes(
  x = Institution,
  y = Count,
  fill = Institution
)) +
  geom_col() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(
    color = "#993333",
    size = 10
  )) +
  ggtitle("Global Perception of corruption by institution, 2013-2017") +
  xlab("Institution") +
  scale_fill_manual(
    values = c(
      "#C9BE4E",
      "#478934",
      "#5CD2FF",
      "#5C7077",
      "#C898DC",
      "#B23F1D",
      "#C25B7E",
      "#AE95CA",
      "#7C213D",
      "#09A2E5"
    )
  ) +
  scale_x_discrete(labels = label_wrap(12))
```

# Visualization 4 & 5

## FATAL ENCOUNTER IN USA, 2015-2022

### Data Sources:

- https://www.washingtonpost.com/graphics/investigations/police-shootings-database/

- https://github.com/washingtonpost/data-police-shootings/tree/master/v1

```{r data_us_plot4,5}
fatal_police_shootings_data <-
  read_csv("data/policingthepolice/fatal-police-shootings-data.csv") |>
  mutate(year = year(date)) |>
  select(gender, race, city, state, year)

shots_usa_plot4 <- fatal_police_shootings_data |>
  group_by(year) |>
  summarise(count = n()) |>
  ggplot(aes(x = year, y = count)) +
  geom_line(colour = "#351873") +
  geom_point() +
  theme_wsj() +
  ggtitle("USA Killings: 2015-2022")
shots_usa_plot4

state_plot5 <- fatal_police_shootings_data |>
  group_by(state) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  head(10) |>
  ggplot(aes(x = state, y = count)) +
  geom_segment(aes(
    x = reorder(state, -count),
    xend = state,
    y = 0,
    yend = count
  )) +
  geom_point(
    size = 6,
    color = "#6F0269",
    fill = alpha("#074B1D", 0.3),
    alpha = 0.7,
    shape = 21,
    stroke = 2
  ) +
  theme_wsj() +
  labs(
    title = "USA States with Highest Killings",
    subtitle = "2015-2022"
  )
state_plot5
```

# Visualization 6 & 7

## REGESTERED CASES AGAINST POLICE VS ARRESTED, 2017-2020

### NCRB 

- https://ncrb.gov.in/en/crime-in-india-table-addtional-table-and-chapter-contents?page=21

```{r data_india_bind}

ncrb_col_names <- c(
  "S. No.",
  "State/UT",
  "Registered Cases",
  "Quashed/ Stayed by Court",
  "Charge-sheeted cases",
  "Final Report Submitted",
  "Arrested",
  " Charge-sheeted police",
  "Cases With-drawn/ Disposed Off",
  "Trials were Completed",
  " Convicted",
  "Acquitted or Discharged",
  "Year")

Table_16A_5_0 <-
  read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5_0.xlsx") |>
  slice(-c(1, 2, 3, 4, 5, 35, 36, 44, 45, 46, 47)) |>
  mutate(Year = "2018") |>
  set_names(ncrb_col_names)

Table_16A_5_1 <-
  read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5_1.xlsx") |>
  slice(-c(1, 2, 3, 4, 5, 35, 36, 44, 45, 46, 47)) |>
  add_column(Year = "2017") |>
  set_names(ncrb_col_names)

Table_16A_5_2 <-
  read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5-2.xlsx") |>
  slice(-c(1, 2, 3, 4, 5, 34, 35, 44, 45, 46, 47)) |>
  add_column(Year = "2020") |>
  set_names(ncrb_col_names)

Table_16A_5 <-
  read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5.xlsx") |>
  slice(-c(1, 2, 3, 4, 5, 35, 36, 44, 45, 46, 47)) |>
  add_column(Year = "2019") |>
  set_names(ncrb_col_names)

data_policecases <-
  bind_rows(Table_16A_5_0, Table_16A_5_1, Table_16A_5_2, Table_16A_5) |>
  mutate(
    Arrested = as.numeric(Arrested),
    `Registered Cases` = as.numeric(`Registered Cases`)
  ) |>
  select(`State/UT`, `Registered Cases`, Arrested, Year) |>
  mutate(
    sum = rowSums(across(where(is.numeric))),
    percentage_arrested = (Arrested / sum),
    `State/UT` = trimws(`State/UT`),
    `State/UT` = recode(`State/UT`, "Delhi UT" = "Delhi")
  )
```
  
```{r plot6}
plot6_percentage <- data_policecases |>
  filter(Year == "2020") |>
  arrange(desc(`Registered Cases`)) |>
  head(15) |>
  mutate(`State/UT` = factor(`State/UT`, `State/UT`)) |>
  ggplot(
    aes(
      x = `Registered Cases`,
      y = percentage_arrested,
      size = `Registered Cases`,
      fill = `State/UT`
    )
  ) +
  geom_point(
    alpha = 0.5,
    shape = 21,
    color = "black"
  ) +
  scale_size(range = c(.1, 24)) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    x = "Number of registered cases",
    y = "Percentage of police arrested out of registered cases",
    size = "No. of police officers in case",
    title = "Top 15 States with highest registered cases, 2020",
  ) +
  theme_classic()

plot6_percentage
```

```{r karnataka_plot7}
plot7_kar <- data_policecases |>
  filter(`State/UT` == "Karnataka") |>
  ggplot() +
  geom_col(
    aes(
      x = reorder(Year, -`Registered Cases`),
      y = `Registered Cases`),
      fill = "#86BCD9") +
  geom_col(aes(x = Year, y = Arrested), fill = "#E95E5E") +
  ggtitle("Registered Cases vs Police Arrested: Karnataka") +
  ylab("Regestered Cases & Arrested (Red") +
  xlab("Year") +
  theme_minimal() +
  annotate("text",
    x = 1,
    y = 10,
    label = "12% Arrested"
  ) +
  annotate("text",
    x = 3,
    y = 8,
    label = "34% Arrested"
  )

plot7_kar
```

# Visualization 8

## EUROPE TRUST IN POLICE (RATING), 2013

### Data and Viz Source:

-OWID: https://ourworldindata.org/grapher/trust-in-others-vs-trust-in-police

```{r plot8}
options(scipen = 999)
europe_trust_in_police <-
  read_csv("data/policingthepolice/trust-in-others-vs-trust-in-police.csv") |>
  na.omit()

ggplot(
  europe_trust_in_police,
  aes(y = `GDP per capita, PPP (constant 2017 international $)`, x = `Trust in police (Eurostat (2015))`)
) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  annotate("text",
    x = 6,
    y = 110000,
    label = "Luxembourg"
  ) +
  annotate("text",
    x = 7.4,
    y = 69877,
    label = "Switzerland"
  ) +
  annotate("text",
    x = 6.4,
    y = 45000,
    label = "UK"
  ) +
  theme_bw()
```


# Visualization 9- ASK

## FINACIAL RESERVE UK 2018

### Data source:

- https://data.police.uk/data/open-data/

```{r plot9}
Police_reserves_data_2018 <-
  read_csv("data/policingthepolice/financial_reserves_data/Police reserves data 2018.csv") |>
  slice(-c(1, 2, 3, 4, 47, 48, 49, 50, 51, 52, 53, 54))
```


# Visualization 10

## PHILLIPPINES DRUG WAR CASUALITIES

### Data source:

-  https://data.world/stabile-center/ph-drug-war#

```{r plot10}
PH_data <- read_excel("data/policingthepolice/PH data.xlsx") |>
  filter(`Type of killing` == "Police killing") |>
  mutate(Date = as.Date(Date, format = "%Y")) |>
  select(City, `Type of killing`, Date) |>
  group_by(City, Date) |>
  summarise(count = n())

ggplot(PH_data, aes(x = Date, y = count, fill = City)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(),
    alpha = 0.75
  ) +
  ggtitle("Phillippines Drug War Police Killings") +
  ylab("People Killed") +
  theme_economist() +
  scale_fill_economist() +
  theme(
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  ) +
  scale_x_date(
    breaks = date_breaks("years"),
    labels = date_format("20%y")
  )
```


# Reference

Basic question for plotting x axis using row.names. (n.d.). Google Groups. https://groups.google.com/g/ggplot2/c/UkmDYDcRNWc?pli=1

Chang, W. (2023, May 28). 7.1 Adding Text Annotations | R Graphics Cookbook, 2nd edition. https://r-graphics.org/recipe-annotate-text

Combine two data frames with the same column names. (n.d.). Stack Overflow. https://stackoverflow.com/questions/20081256/combine-two-data-frames-with-the-same-column-names

Crime in India Table Contents |  National Crime Records Bureau. (n.d.). https://ncrb.gov.in/en/crime-in-india-table-addtional-table-and-chapter-contents?field_date_value%5Bvalue%5D%5Byear%5D=&field_select_table_title_of_crim_value=18&items_per_page=All

Formatting Decimal places in R. (n.d.). Stack Overflow. https://stackoverflow.com/questions/3443687/formatting-decimal-places-in-r

GeeksforGeeks. (2021). Change column name of a given DataFrame in R. GeeksforGeeks. https://www.geeksforgeeks.org/change-column-name-of-a-given-dataframe-in-r/

ggplot2 axis scales and transformations - Easy Guides - Wiki - STHDA. (n.d.). http://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations

ggplot2 scatter plots : Quick start guide - R software and data visualization - Easy Guides - Wiki - STHDA. (n.d.). http://www.sthda.com/english/wiki/ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization

Highlight a single “bar” in ggplot. (n.d.). Stack Overflow. https://stackoverflow.com/questions/45820250/highlight-a-single-bar-in-ggplot

how to add lines over a column bar graph where the lines pass by the middle-top of the bars considering bars with position=’dodge’? (n.d.). Stack Overflow. https://stackoverflow.com/questions/72116660/how-to-add-lines-over-a-column-bar-graph-where-the-lines-pass-by-the-middle-top

Plot data in descending order as appears in data frame. (n.d.). Stack Overflow. https://stackoverflow.com/questions/16961921/plot-data-in-descending-order-as-appears-in-data-frame

Side By Side Bar Graphs In R & ggplot2. (n.d.). https://dk81.github.io/dkmathstats_site/rvisual-sidebyside-bar.html

Sum across multiple columns with dplyr. (n.d.). Stack Overflow. https://stackoverflow.com/questions/28873057/sum-across-multiple-columns-with-dplyr