---
title: "Policing the police"
author: "Nishtha Arora"
date: "2023-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(readr)
library(readxl)
library(tidyverse)
```

Potential data sets

# UNODC

```{r}
corruption_UNODC <- read_excel("data/policingthepolice/data_cts_corruption_and_economic_crime.xlsx") |> slice(-c(1,2))
colnames(corruption_UNODC) <- c('Iso3_code','Country','Region', 'Subregion','Indicator','Dimension', 'Category','Sex','Age','Year','Unit of measurement','VALUE','Source') 
```

#OWID

Q calculate the % of corruption done by police in 2017.
Perception of corruption by institution, World, 2017 and Corruption Perception Index, 2012 to 2018- https://ourworldindata.org/corruption
```{r}
OWID_police <- read_csv("data/policingthepolice/perception-of-corruption-by-institution.csv")

TI_corruption_perception_index <- read_csv("data/policingthepolice/TI-corruption-perception-index.csv") |> 
  filter(Year == "2017")
```

# Washington Post/Github

Fatal Encounters-USA
https://www.washingtonpost.com/graphics/investigations/police-shootings-database/
https://github.com/washingtonpost/data-police-shootings/tree/master/v1
```{r}
fatal_police_shootings_data <- read_csv("data/policingthepolice/fatal-police-shootings-data.csv")
```


# NCRB 

Cases against police-India
https://ncrb.gov.in/en/crime-in-india-table-addtional-table-and-chapter-contents?page=21
```{r}
Table_16A_5_0 <- read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5_0.xlsx") |> 
  slice(-c(1,2,3,4,5, 35, 36, 44, 45, 46,47)) |> 
  add_column(Year = "2018")
colnames(Table_16A_5_0) <- c('S. No.','State/UT','Registered Cases', 'Quashed/ Stayed by Court','Charge-sheeted cases','Final Report Submitted', 'Arrested',' Charge-sheeted police','Cases With-drawn/ Disposed Off','Trials were Completed',' Convicted','Acquitted or Discharged', 'Year') 

Table_16A_5_1 <- read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5_1.xlsx") |> 
  slice(-c(1,2,3,4,5, 35, 36, 44, 45, 46,47)) |> 
  add_column(Year = "2017")
colnames(Table_16A_5_1) <- c('S. No.','State/UT','Registered Cases', 'Quashed/ Stayed by Court','Charge-sheeted cases','Final Report Submitted', 'Arrested',' Charge-sheeted police','Cases With-drawn/ Disposed Off','Trials were Completed',' Convicted','Acquitted or Discharged', 'Year') 

Table_16A_5_2 <- read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5-2.xlsx") |> 
  slice(-c(1,2,3,4,5, 35, 36, 44, 45, 46,47)) |> 
  add_column(Year = "2020")
colnames(Table_16A_5_2) <- c('S. No.','State/UT','Registered Cases', 'Quashed/ Stayed by Court','Charge-sheeted cases','Final Report Submitted', 'Arrested',' Charge-sheeted police','Cases With-drawn/ Disposed Off','Trials were Completed',' Convicted','Acquitted or Discharged', 'Year') 


Table_16A_5 <- read_excel("data/policingthepolice/castagainstpolicep/Table 16A.5.xlsx")|> 
    slice(-c(1,2,3,4,5, 35, 36, 44, 45, 46,47)) |> 
  add_column(Year = "2019")
colnames(Table_16A_5) <- c('S. No.','State/UT','Registered Cases', 'Quashed/ Stayed by Court','Charge-sheeted cases','Final Report Submitted', 'Arrested',' Charge-sheeted police','Cases With-drawn/ Disposed Off','Trials were Completed',' Convicted','Acquitted or Discharged', 'Year') 

data_policecases <- rbind(Table_16A_5_0, Table_16A_5_1, Table_16A_5_2, Table_16A_5)
```


# Reference

Combine two data frames with the same column names. (n.d.). Stack Overflow. https://stackoverflow.com/questions/20081256/combine-two-data-frames-with-the-same-column-names

Crime in India Table Contents |  National Crime Records Bureau. (n.d.). https://ncrb.gov.in/en/crime-in-india-table-addtional-table-and-chapter-contents?field_date_value%5Bvalue%5D%5Byear%5D=&field_select_table_title_of_crim_value=18&items_per_page=All

GeeksforGeeks. (2021). Change column name of a given DataFrame in R. GeeksforGeeks. https://www.geeksforgeeks.org/change-column-name-of-a-given-dataframe-in-r/