---
title: "**DATA VISUALIZATION PROJECT**: Understanding trends, outliers and patterns in Data."
subtitle: "ETC5543- S1 2023"
author: "Nishtha Arora"
date: "2023-06-15"
output:
  bookdown::html_document2:
        base_format: rmdformats::readthedown
        keep_md: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height = 6, fig.width = 4)
```

```{r}
library(rmdformats)
```

# Abstract {-}

This document provides the discussion on visualization process for selected publishing articles; Decrimianlizing Suicide and Policing the Police, for the organization 360info.org. Details such as, data sources, methodology, findings, significance and motivation for analysis are reported.

# About  {-}

something

# Publication: Decrimanlizing Suicide  

## Background 

## Motivation and Significance

## Objectives

## Methodology

### Data

### Results

### Discussion

# Publication: Policing the Police

## Background 

## Motivation and Significance

## Objectives

## Methodology

### Data

### Results 

### Discussion

# Conclusion  {-}

# References  {-}





