---
title: "Data Visualization Need Analysis: The Design Process"
subtitle: "ETC5543 Business Analysis Creative Activity, Semester 1-2023"
author: "Nishtha Arora (32296622)"
institute: " Monash University, Clayton"
date: "March 2023 - May 2023"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: style.css
    yolo: FALSE
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(themes360info)
library(readxl)
library(tidyverse)
library(hrbrthemes)
library(ggthemes)
library(lubridate)
library(sf)
library(rgeos)
library(pdftools)
library(ggthemes)
library(scales)
library(dumbbell)
```

class: center, middle

# DECRIMANALIZING SUICIDE

- Worldwide Public Health Problem

- over 700,000 deaths from suicide worldwide in 2019

- decriminalizing suicide is a boon or a bane?

- still 20 countries that have criminalized suicide

## History

With the advancement of science in the 19th and 20th century, it was discovered that suicidal tendencies are caused by biological factors as well and hence, continents like Europe and North America revoked the laws reagrding crimanlizng suicide.

---

### Visualization 1

```{r saving_cleaned_data, echo=FALSE, message=FALSE, warning=FALSE}
suicide_hdi_data <- read_csv("data/join_HDI_suicide.csv")

plot1 <- suicide_hdi_data %>% head(30)
ggplot(plot1,
  aes(x = reorder(Entity, -avg_rate), y = avg_rate,
    fill = `Development status`)) +
  geom_col() +
  geom_text(aes(label = avg_rate), vjust = 2, colour = "white", size = 1.5) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    plot.title = element_text(face = "bold")
  ) +
  labs(
    title = "Top 30 countries by average suicide rate (2008-2019)",
    x = substitute(paste(bold("Country"))),
    y = substitute(paste(bold("Suicide rate per 100,000 people")))
  ) +
  scale_fill_manual(
    values = c("#A0331C", "#1C56A0", "#4B902F", "#635A61"))
```

---

### Visualization 2

```{r who_data_plot2, echo=FALSE, message=FALSE, warning=FALSE}
who_data <- read_csv("data/data-2.csv") |>
  filter(Dim1 == "Both sexes") |>
  select(SpatialDimValueCode, Location, Period, Dim1, FactValueNumeric)

who_data |> filter(Location == "India") |>
  ggplot(aes(x = Period, y = FactValueNumeric)) +
  geom_line() +
  geom_point(color = "brown", size = 1.5) +
  geom_vline(xintercept = 2017, linetype = "dashed") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2019, by = 1)) +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold"),
    plot.title = element_text(size = 14, lineheight = .8)
  ) +
  xlab(NULL) +
  ylab(substitute(paste(bold(
    "Rate per 100,000 people"
  )))) +
  ggtitle("Time-series for India, 2000-2019") +
  theme(plot.title = element_text(size = 14, lineheight = .8)) +
  geom_vline(xintercept = 2017, linetype = "dashed") +
  scale_color_manual(values = c('Brown', 'Purple', 'Black')) +
  theme(legend.position = "none")
```

---

### Visualization 3


